Evaluative adjective sentences: \\ A question-based analysis of projection

### 2019-07-05 paper/paper-anon.tex
updated paper and response letter

@@

Two contents of evaluative adjective sentences, like {\em Kim was smart to watch the movie}, are the prejacent (that Kim watched the movie) and the generalization (that the degree to which Kim watching the movie is smart was higher than the contextual standard of {\em smart}). The prejacent is standardly analyzed as a presupposition
(e.g., \citealt{norrick78,barker02,oshima09b,kertz2010}). This paper argues against such analyses of the prejacent because they, among other things, do not capture an interaction between the prejacent and the generalization that has not yet been observed for projective content: when the prejacent projects, the generalization does not, and when the prejacent does not project, the generalization does. We develop an analysis according to which the prejacent is not a lexically specified presupposition but projects if it is not at-issue with respect to the question addressed by the utterance of the evaluative adjective sentence. In addition to capturing the interaction between the prejacent and the generalization, our question-based projection analysis significantly extends previous such analyses (e.g., \citealt{beaver-clark08,best-question,brst-ar}) by incorporating a novel constraint on the question addressed by an utterance: the more the truth of an utterance content is taken to follow from the common ground, the less likely the question is about that content. We provide experimental evidence for the proposed analysis and argue that our analysis improves on that of \citealt{karttunen-etal2014}, according to which evaluative adjectives are systematically ambiguous.

@@

In an evaluative adjective sentence (EAS) like (\ref{f}), the
evaluative adjective {\em stupid} subcategorizes for a non-pleonastic subject noun
phrase ({\em Feynman}) and a {\em to-}infinitive ({\em to
dance on the table}); the subject of the predicate {\em be stupid} is the understood subject of the {\em to-}infinitive (e.g., \citealt{wilkinson70, norrick78,barker02,kertz2010}). One of the contents standardly discussed in the literature is (what we refer to as) the prejacent: in (\ref{f}), the prejacent is that Feynman danced on the table.

@@

Recently, \citet{karttunen-etal2014} provided naturally occurring examples that show that utterances of sentences in which the evaluative adjective is embedded under negation can receive an interpretation according to which the prejacent does not project, i.e., can be interpreted in the scope of negation. For instance, the speaker of (\ref{nat}a) is not committed to living close to their parents and instead communicates that they do not live close to their parents. Likewise, the speaker of (\ref{nat}b) communicates that they did not go stumbling through the junkyard and get hurt.\footnote{Non-projection of the prejacent is also observed with  entailment-canceling operators, such as the antecedent of a conditional in B's utterance in (ia) or the possibility modal {\em perhaps} in (ib). While this paper limits its attention to EAS embedded under negation, we expect the analysis to carry over to other entailment-canceling operators.

@@

Before discussing why such examples are problematic for analyses of the prejacent as a lexically specified presupposition, we would first like to acknowledge that there is variation in the population of native speakers of American English: whereas speakers are generally able to interpret EASs in which the prejacent does not project, i.e., are able to retrieve the intended interpretations of examples like (\ref{nat}), a sizeable portion strongly prefers to realize such interpretations with variants that includes {\em enough}, as in {\em I wasn't stupid enough to go stumbling through the junkyard in the dark} for (\ref{nat}b); see also \citealt{karttunen2013} and \citealt{karttunen-etal2014} for this observation. Nevertheless, EASs in which the prejacent does not project are part of American English, as evidenced by the existence of naturally occurring examples like (\ref{nat}). Additional evidence comes from the fact that there are native speakers who judge such EASs to be perfectly acceptable\footnote{The audiences to which this research was presented over the years always included native speakers that judged examples like (\ref{nat}) to be acceptable. More systematic evidence for the existence of such speakers comes from the acceptability rating study presented in Appendix \ref{s-acc}: of the 94 self-reported native speakers of American English that participated in the study, about 20-30\% judged negated evaluative adjective sentences without {\em enough} to be acceptable under a non-projecting interpretation of the prejacent.} and who produce examples like (\ref{nat}). We therefore assume that native speakers of American English are generally able to retrieve both interpretations of EASs even if they might not produce EASs in which the prejacent does not project. Our goal in this paper is to analyze the interpretation of EASs; we briefly return to the observed variation after developing our analysis.

To account for examples like (\ref{nat}), in which the prejacent does not project, presupposition analyses appeal to local accommodation, a process whereby a presupposition is added to a local context, such as that created by negation: presuppositions can be locally accommodated if adding the presupposition to the common ground of the interlocutors (a process known as global accommodation) would result in a contradiction, uninformativity or problems with
binding (\citealt{heim82,vds92}). In example (\ref{nat}a), for instance, the prejacent, that the speaker lived close to their parents, is locally accommodated under negation because globally accommodating the prejacent in the common ground would result in a contradiction: 
according to the context, the closest that the speaker lived to their parents was when the speaker lived in Denver, Colorado, i.e., about 300 miles (480 km) away from their parents in Garden City, Kansas. The prejacent is correctly predicted to be locally accommodated in (\ref{nat}a) and thereby is not a commitment of the speaker.

Negated evaluative adjective sentences (NEASs) like (\ref{nat}) point to two problems for analyses according to which the prejacent is a presupposition. The first problem concerns the prejacent. In example (\ref{nat}b), globally accommodating the prejacent does not result in a contradiction, uninformativity or problems with binding: if the prejacent, that the speaker went stumbling in the junkyard, was added to the common ground, (\ref{nat}b) would mean that the speaker knew that someone was in the junkyard, that they stumbled through the junkyard and got hurt, and that the speaker does not consider these actions stupid (perhaps because these actions led to the person in the junkyard receiving help). Because such an interpretation does not result in a contradiction, uninformativity or problems with binding, analyses of the prejacent as a presupposition incorrectly  predict that the prejacent of (\ref{nat}b) is globally accommodated, i.e., is a commitment of the speaker.\footnote{It may be possible to account for the non-projection of the prejacent on the basis of plausibility considerations: \citealt[162]{vonfintel08}, for instance,  proposed that ``what gets accommodated depends on the best guess of the hearers about what the speaker might have intended as the adjustment to the common ground that would admit the asserted sentence''. The cues to the projectivity of the prejacent identified in the following sections may help flesh out the reasoning process that hearers and readers undergo in interpreting EASs. Plausibility considerations do not, however, address the second problem that presuppositional analyses face.}

@@

The examples in (\ref{nat}) reveal a remarkable interaction that has not yet been observed for projective content and that was not noted in \citealt{karttunen-etal2014}: when the prejacent projects, the generalization does not, as in (\ref{f2}b), and when the prejacent does not project, the generalization does, as in (\ref{nat}b). To illustrate that this behavior is strikingly different from the interpretation of utterances of other sentences that give rise to projective content consider  (\ref{stop}). When the content of the clausal complement of {\em know}, that the meeting was canceled, projects, what is denied is Sam's knowledge of this content. The critical difference between EASs and examples like (\ref{stop}) comes out when the content of the complement does not project, as is brought out, for instance, by continuing (\ref{stop}) with {\em \ldots he, like all of us, is in the dark about whether the meeting will take place}. In this case, Sam's knowledge of the content that the meeting was canceled is still denied and the speaker is not taken to be committed to either its truth or its falsity. Thus, the two contents of (\ref{stop}) do not exhibit the interaction of the projectivity of the prejacent and the generalization.

\begin{exe}
\ex\label{stop} Sam doesn't know that the meeting was canceled.
\end{exe}
This observation has strong implications for analyses of EASs according to which the prejacent is a lexically specified presupposition. First, when the prejacent is locally accommodated under negation, the speaker is committed to the falsity of the prejacent; this is in contrast to a locally accommodated factive presupposition, for which the speaker is not committed to its truth or falsity. Second, when the prejacent is locally accommodated under some operator, the generalization is not interpreted under that operator, in contrast to the attitude ascription with {\em know} which is always interpreted in the scope of the operator. These observations are problematic for advocates of analyses according to which the prejacent is a lexically specified presupposition and projection is assumed to be governed by the standard mechanisms of presupposition projection (e.g., \citealt{heim83,vds92}).

\citepos{karttunen-etal2014} analysis of EASs made progress over previous analyses by assuming that evaluative adjectives are ambiguous between the two schematic lexical entries in (\ref{lex}): the prejacent is specified as presupposed in the lexical entry in  (\ref{lex}a) but not in (\ref{lex}b).

\begin{exe}
\ex\label{lex} \citealt[249]{karttunen-etal2014}: Presuppositions and assertions of EASs of the form `NP was Adj to VP'
\begin{xlist}
\ex {\em adj}$ 1$
\\ Presupposed content: NP VPed
\\ Asserted content: For NP to VP would be Adj

\ex {\em adj}$ 2$
\\ Presupposed content: For NP to VP would be Adj \& for NP not to VP would not be Adj
\\ Asserted content: What NP did about VPing was Adj

@@

In the next section, we develop an analysis of EASs under which evaluative adjectives only have a single lexical entry. The analysis correctly predicts the two attested interpretations of EASs and the  interaction between the projection of the prejacent and the generalization falls out of the analysis. {\bf ADD PARAGRAPH HERE ABOUT DIFFERENT STYLES OF PROJECTION ANALYSES, THAT PREJACENT AND GENERALIZATION WILL BE ATTRIBUTED TO DISCOURSE, BUT THAT OTHER CONTENT IS PROJECTIVE BECAUSE OF CONVENTIONAL SPECIFICATION. \citealt{kadmon01,brst-lang11,tonhauser-etal-sub23}} Section \ref{s3} provides experimental evidence for two predictions of the analysis. After briefly considering the aforementioned interspeaker variation in section \ref{s5}, the paper concludes in section \ref{s6} {\bf DISCUSS HERE AGAIN THAT CONTENT PROJECTS FOR DIFFERENT REASONS}

As discussed above, previous analyses of EASs have attributed the projectivity of the prejacent to its lexical specification as a presupposition. More recently, it has been observed that some utterance content is not highly projective and that its projection is sensitive to the discourse context, and these observations have led to the development of analyses according to which projection is not derived from a lexical specification but rather derived from projective content being backgrounded or not-at-issue (e.g., \citealt{abrusan2011,abrusan2016,brst-salt10,best-question,brst-ar}). In this paper, we follow such proposals and argue that the prejacent of an EAS projects when it is not at-issue with respect to the question addressed by the utterance of the EAS. Empirical motivation for developing such an analysis comes from the observation that the prejacent of EASs is not highly projective.  \citepos{tbd-variability} experimental investigation found that the projectivity of the prejacent of EASs with {\em stupid} was significantly lower than that of non-restrictive relative clauses and appositive content, as well as of the content of the complement of the factive predicates {\em be annoyed} and {\em know}. In naturally occurring data, too, the prejacent is not highly projective. In a corpus-based web study, we collected `certainty' ratings on a 7-point Likert scale from 226 native speakers of American English for the prejacents of 59 naturally occurring NEASs: the higher the certainty rating, the more projective the prejacent. Figure \ref{f-corpus} shows the mean certainty ratings of the 59 NEASs by evaluative adjective: across the examples, the prejacent is not highly projective; the mean certainty rating was only 3.2. Details on this study are given in Appendix \ref{a-corpus}.

@@

In this section, we develop a question-based projection analysis for EASs.\footnote{This is not to suggest that lexicalist projection analyses are never appropriate: we assume, for instance, that contents associated with what \citealt{brst-lang11} refer to as a Strong Contextual Felicity constraint are lexically specified presuppositions and that the projection of conventional implicatures is best attributed to conventional specification (e.g., \citealt{potts05,murray2014}). For discussion  see \citealt{brst-lang11,tbd-variability,tonhauser-etal-sub23}. The prejacent of sentences like {\em It was smart of Kim to watch the movie} may also be best analyzed as a lexically specified presupposition if, as native speaker intuitions suggest, it is indeed highly projective and not-at-issue, in contrast to the prejacent of EASs.} We start in section \ref{s21} by introducing relevant aspects of the question-based projection analysis developed in \citealt{best-question}. In section \ref{s22}, we show that the prejacent and the generalization are lexical entailments of EASs and that, furthermore, exactly one of them is at-issue in any given utterance; this predicts the interaction in the projection of these two contents. In section \ref{s23}, we extend previous question-based analyses by incorporating a constraint on the questions addressed by utterances of EASs.

@@

As illustrated in section \ref{s1}, it is a long-standing observation that the content of the complement of an example like (\ref{stop}) {\em Sam doesn't know that the meeting was canceled} can project, but need not. Under lexicalist projection analyses, the projection of the content of the complement is predicted by specifying it as a presupposition in the lexical entry of the factive predicate (e.g., \citealt{heim83,vds92,schlenker10}. Several non-lexicalist analyses have been proposed over the years: these argue that the projection of the content of the complement is due to pragmatic considerations in order to capture, among other things, that content of the complement is not highly projective and that its projection is sensitive to the discourse context (e.g., \citealt{wilson75,simons01,abusch02,abusch10,abrusan2011,abrusan2016,romoli2015}). In this paper, we adopt \citepos{best-question} question-based projection analysis for the content of the complement of factive predicates to account for the projective content of EASs.

According to \citealt{best-question}, one of the conditions under which the content of the complement of a factive predicate projects, i.e., is understood as a commitment of the speaker, is if the content of the complement is not at-issue with respect to the Discourse Question addressed by the utterance of the sentence with the factive predicate. The Discourse Question ``provides the topic of a segment of discourse and imposes relevance constraints on conversational contributions'' (p.192) and utterance content is at-issue with respect to the Discourse Question of the utterance if the content addresses the Discourse Question, i.e., entails at least a partial answer to the question (\citealt{roberts12}/1996). To illustrate, consider the content of the clausal complement of {\em discover} in B's utterances in (\ref{discover2}a) and (\ref{discover2}b), that Harriet was at Princeton for a job interview. In the these examples, the Discourse Questions that B's utterances address are made explicit by A's interrogative utterances. In (\ref{discover2}a), the content of the  complement of {\em discover} is at-issue because it addresses the Discourse Question: that Harriet was at Princeton for a job interview is an answer A's interrogative utterance of where Harriet was yesterday. In (\ref{discover2}b), on the other hand, the content of the complement is not at-issue because it is not an answer to A's interrogative utterance; rather, here the Discourse Question is addressed by the main clause content of B's utterance.

@@

Now consider the example in (\ref{discover22}), in which  {\em discover} is embedded under an entailment-canceling operator. Here, the speaker is not taken to be committed to the content of the main clause; rather, the speaker is asking whether Henry discovered the content of the complement.

\begin{exe}
\ex\label{discover22} Did Henry discover that Harriet was at Princeton for a job interview?
\end{exe}
In contrast to the main clause content, the content of the complement of {\em discover} is not merely an entailment but may project, as noted above.\footnote{For the assumption that at least some presuppositions, as well as other projective content, are entailments that may project see, e.g., \citealt{gazdar79b,barker02,schlenker10,abrusan2011,abrusan2016,anand-hacquard2014}. } According to \citealt{best-question}, one of the conditions under which the content of the complement projects is when it is not at-issue with respect to the Discourse Question addressed by the utterance. Consider the examples in (\ref{discover3}), where A's interrogative utterances again make explicit the Discourse Questions addressed by B's utterances. In (\ref{discover3}a), the content of the complement of B's utterance does not address the Discourse Question, i.e., is not at-issue. Here, B is taken to be committed to the content of the complement: to make sense of B's utterance as an answer to A's question, there must be a connection between the possibility of Henry discovering something about Harriet and Henry's mood; an obvious connection is that B takes the content of the complement to be true, i.e., that it is possible that Henry discovered that Harriet was at Princeton for a job interview. Thus, the content of the complement projects. In (\ref{discover3}b), by contrast, the content of the complement addresses the Discourse Question, i.e., is at-issue. Here, B need not be taken to be committed to the content of the complement to make sense of how B's utterance addresses A's question and the content of the complement does not project.

@@

\end{exe}} For empirical evidence that at-issueness predicts projection, see \citealt{tbd-variability}, who investigated the hypothesis for 19 projective contents associated with American English expressions.

>>>>

\vspace*{-.7cm}

@@

Two contents of evaluative adjective sentences, like {\em Kim was smart to watch the movie}, are the prejacent (that Kim watched the movie) and the generalization (that the degree to which Kim watching the movie is smart was higher than the contextual standard of {\em smart}). The prejacent is standardly analyzed as a presupposition
(e.g., \citealt{norrick78,barker02,oshima09b,kertz2010}). This paper argues against such analyses of the prejacent because, among other things, they do not capture an interaction between the prejacent and the generalization that has not yet been observed for projective content: when the prejacent projects, the generalization does not, and when the prejacent does not project, the generalization does. We develop an analysis according to which the prejacent is not a lexically specified presupposition but projects if it is not at-issue with respect to the question addressed by the utterance of the evaluative adjective sentence. In addition to capturing the interaction between the prejacent and the generalization, our question-based projection analysis significantly extends previous such analyses (e.g., \citealt{beaver-clark08,best-question,brst-ar}) by incorporating a novel constraint on the question addressed by an utterance: the more the truth of an utterance content is taken to follow from the common ground, the less likely the question is about that content. We provide experimental evidence for the proposed analysis and argue that our analysis improves on that of \citealt{karttunen-etal2014}, according to which evaluative adjectives are systematically ambiguous.

@@

In an evaluative adjective sentence (EAS) like (\ref{f}), an
evaluative adjective\footnote{Evaluative adjectives are adjectives like {\em stupid, rude} or {\em fortunate} that, in sentences like {\em Sam was stupid/rude/fortunate}, convey the speaker's positive or negative evaluation of or attitude towards the denotation of the subject noun phrase.} ({\em stupid}) subcategorizes for a non-pleonastic subject noun
phrase ({\em Feynman}) and a {\em to-}infinitive ({\em to
dance on the table}); the subject of the predicate ({\em be stupid}) is the understood subject of the {\em to-}infinitive (e.g., \citealt{wilkinson70, norrick78,barker02,kertz2010}).  One of the contents standardly discussed in the literature is (what we refer to as) the prejacent: in (\ref{f}), the prejacent is that Feynman danced on the table.

@@

Recently, \citet{karttunen-etal2014} provided naturally occurring examples that show that utterances of sentences in which the evaluative adjective is embedded under negation can receive an interpretation according to which the prejacent does not project, i.e., can be interpreted in the scope of negation. For instance, the speaker of (\ref{nat}a) is not committed to living close to their parents and instead communicates that they do not live close to their parents. Likewise, the speaker of (\ref{nat}b) communicates that they did not go stumbling through the junkyard and get hurt.\footnote{Non-projection of the prejacent is also observed with  other entailment-canceling operators, such as the antecedent of a conditional in B's utterance in (ia) or the possibility modal {\em perhaps} in (ib). While this paper limits its attention to EASs embedded under negation, we expect the analysis to generalize  to other entailment-canceling operators.

@@

Before discussing why such examples are problematic for analyses of the prejacent as a lexically specified presupposition, we would first like to acknowledge that there is variation in the population of native speakers of American English: whereas speakers are generally able to interpret EASs in which the prejacent does not project, i.e., are able to retrieve the intended interpretations of examples like (\ref{nat}), a sizeable portion strongly prefers to realize such interpretations with variants that include {\em enough}, as in {\em I wasn't stupid enough to go stumbling through the junkyard in the dark} for (\ref{nat}b); see also \citealt{karttunen2013} and \citealt{karttunen-etal2014} for this observation. Nevertheless, EASs in which the prejacent does not project are part of American English, as evidenced by the existence of naturally occurring examples like (\ref{nat}). Additional evidence comes from the fact that there are native speakers who judge such EASs to be perfectly acceptable\footnote{The audiences to which this research was presented over the years always included native speakers that judged examples like (\ref{nat}) to be acceptable. More systematic evidence for the existence of such speakers comes from the acceptability rating study presented in Appendix \ref{s-acc}: of the 94 self-reported native speakers of American English that participated in the study, about 20-30\% judged negated evaluative adjective sentences without {\em enough} to be acceptable under a non-projecting interpretation of the prejacent.} and who produce examples like (\ref{nat}). We therefore assume that native speakers of American English are generally able to retrieve both interpretations of EASs even if they might not produce EASs in which the prejacent does not project. Our goal in this paper is to analyze the interpretation of EASs; we briefly return to the observed production variation after developing our analysis.

To account for examples like (\ref{nat}), in which the prejacent does not project, presupposition analyses appeal to local accommodation, a process whereby a presupposition is added to a local context, such as that created by negation: presuppositions can be locally accommodated if the default global accommodation -- adding the presupposition to the common ground of the interlocutors -- would result in a contradiction, uninformativity or problems with
binding (\citealt{heim82,vds92}). In example (\ref{nat}a), for instance, the prejacent, that the speaker lived close to their parents, is locally accommodated under negation because global accommodation would result in a contradiction: 
according to the context, the closest that the speaker lived to their parents was when the speaker lived in Denver, Colorado, i.e., about 300 miles (480 km) away from their parents in Garden City, Kansas. The prejacent is correctly predicted to be locally accommodated in (\ref{nat}a) and thereby is not a commitment of the speaker.

Negated evaluative adjective sentences (NEASs) like (\ref{nat}) point to two problems for analyses according to which the prejacent is a presupposition. The first problem concerns the prejacent. In example (\ref{nat}b), globally accommodating the prejacent does not result in a contradiction, uninformativity or problems with binding: if the prejacent, that the speaker went stumbling in the junkyard, was added to the common ground, (\ref{nat}b) would mean that the speaker knew that someone was in the junkyard, that they stumbled through the junkyard and got hurt, and that the speaker does not consider these actions stupid (perhaps because these actions led to the person in the junkyard receiving help). Because such an interpretation does not result in a contradiction, uninformativity or problems with binding, analyses of the prejacent as a presupposition incorrectly  predict that the prejacent of (\ref{nat}b) is globally accommodated, i.e., is a commitment of the speaker.\footnote{It may be possible to account for the non-projection of the prejacent on the basis of plausibility considerations: \citealt[162]{vonfintel08}, for instance,  proposed that ``what gets accommodated depends on the best guess of the listeners about what the speaker might have intended as the adjustment to the common ground that would admit the asserted sentence''. The cues to the projection of the prejacent identified in this paper may help flesh out the reasoning process that listeners/readers undergo in interpreting EASs. Plausibility considerations do not, however, address the second problem that presuppositional analyses face.}

@@

The examples in (\ref{f2}) and (\ref{nat}) reveal a remarkable interaction that has not yet been observed for projective content and that was not noted in \citealt{karttunen-etal2014}: when the prejacent projects, the generalization does not, as in (\ref{f2}b), and when the prejacent does not project, the generalization does, as in (\ref{nat}b). To illustrate that this behavior is strikingly different from the interpretation of utterances of other sentences that give rise to projective content consider  (\ref{stop}). When the content of the complement of {\em know}, that the meeting was canceled, projects, what is denied is Sam's knowledge of this content. The critical difference between EASs and examples like (\ref{stop}) comes out when the content of the complement does not project, as is brought out, for instance, by continuing (\ref{stop}) with {\em \ldots he, like all of us, is in the dark about whether the meeting will take place}. In this case, Sam's knowledge of the content that the meeting was canceled is still denied and the speaker is not taken to be committed to either its truth or its falsity. Thus, the two contents of (\ref{stop}) do not exhibit the interaction of the projectivity of the prejacent and the generalization.

\begin{exe}
\ex\label{stop} Sam doesn't know that the meeting was canceled.
\end{exe}
This observation has strong implications for analyses of EASs according to which the prejacent is a lexically specified presupposition. First, when the prejacent is locally accommodated under negation, the speaker is committed to the falsity of the prejacent; this is in contrast to a locally accommodated factive presupposition, for which the speaker is not committed to its truth or falsity. Second, when the prejacent is locally accommodated under some operator, the generalization is not interpreted under that operator, in contrast to the attitude ascription with {\em know}, which is always interpreted in the scope of the operator. These observations are problematic for advocates of analyses according to which the prejacent is a lexically specified presupposition and projection is assumed to be governed by the standard mechanisms of presupposition projection (e.g., \citealt{heim83,vds92}).

\citepos{karttunen-etal2014} analysis of EASs made progress over previous analyses; it did so by assuming that evaluative adjectives are ambiguous between the two schematic lexical entries in (\ref{lex}): the prejacent is specified as presupposed in the lexical entry {\em adj}$_1$ in  (\ref{lex}a) but not in that of {\em adj}$_2$ in (\ref{lex}b).

\begin{exe}
\ex\label{lex} \citealt[249]{karttunen-etal2014}: Presuppositions and assertions of EASs of the form `NP was Adj to VP'
\begin{xlist}
\ex {\em adj}$_1$
\\ Presupposed content: NP VPed
\\ Asserted content: For NP to VP would be Adj

\ex {\em adj}$_2$
\\ Presupposed content: For NP to VP would be Adj \& for NP not to VP would not be Adj
\\ Asserted content: What NP did about VPing was Adj

@@

In this paper, we develop an analysis that derives the two interpretations of EASs and the interaction between the projection of the prejacent and the generalization from a single lexical entry for evaluative adjectives. This is achieved, in a nutshell, by making the projection of the prejacent and the generalization sensitive to their discourse status and not lexically specifying them as presupposed. In doing so, our analysis builds on previous analyses according to which projection of content is derived from its discourse status, such as being backgrounded or not-at-issue (e.g., \citealt{abrusan2011,abrusan2013,abrusan2016,brst-salt10,best-question,brst-ar,tbd-variability,tonhauser-etal-sub23}). Analyses in this tradition have been developed for utterance content whose projectivity is sensitive to the discourse context, such as the prejacent of manner adverbs (e.g., that Sam ran in {\em Sam ran quickly}), the pre-state content of {\em stop} (e.g., that Sam whistled in {\em Sam stopped whistling}) or the content of the complement of factive predicates like {\em know} in (\ref{stop}). Such analyses predict not only that the projectivity of such content is sensitive to its discourse status, but also correctly predict that it is less projective than content whose projection is conventionally specified, like the content of non-restrictive relative clauses or the projective content of anaphoric expressions like pronouns or {\em too} (for discussion see, e.g., \citealt{kadmon01,potts05,abrusan2011,abrusan2016,brst-lang11,tbd-variability,tonhauser-etal-sub23}). What it means for one content to be less projective than another is that listeners are less likely to take the speaker to be committed to the truth of the first content than the second. To illustrate, consider the examples in (\ref{proj-var}), in which the content of the complement of {\em know} in (\ref{proj-var}a) and the content of the non-restrictive relative clause in (\ref{proj-var}b) is that the meeting was canceled. Both contents are projective: a speaker who utters the first sentences of these examples may be taken to be committed to the meeting having been canceled. However, listeners are less likely to take the speaker of (\ref{proj-var}a) to be committed to the meeting having been cancelled than the speaker of (\ref{proj-var}b): evidence for this difference in projection comes from the fact that the second sentence is an acceptable continuation in (\ref{proj-var}a) but not (\ref{proj-var}b). For experimental evidence that projection is not a binary, categorical property of utterance content but rather a gradient one see \citealt{smith-hall11,xue-onea11} and \citealt{tbd-variability,tonhauser-etal-sub23}.\footnote{According to \citealt[498f.]{tbd-variability}, there are at least two interpretations of what it means for projectivity to be a gradient rather than a categorical, binary property of utterance content:

Like \citealt{tbd-variability}, we remain agnostic about the underlying interpretation of projectivity as a gradient property, though our discussion of projection variability will be in line with the first interpretation.}

\ex\label{proj-var}

\ex Sam doesn't know that the meeting was canceled. He, like all of us, is in the dark about whether the meeting will take place.

\ex It's not the case that the meeting, which was canceled, was going to take place in the conference room. \#I am in the dark about whether the meeting will take place.

Empirical motivation for adopting an analysis of the prejacent of EASs according to which its projectivity is not derived from its status as a lexically specified presupposition but from its discourse status comes from the observation that the prejacent of EASs is not highly projective.  \citepos{tbd-variability} experimental investigation found that the projectivity of the prejacent of EASs with {\em stupid} was significantly lower than that of non-restrictive relative clauses and appositive content, as well as of the content of the complement of the factive predicates {\em be annoyed} and {\em know}. In naturally occurring data, too, the prejacent is not highly projective: in a corpus-based web study, we collected `certainty' ratings on a 7-point Likert scale from 226 native speakers of American English for the prejacents of 59 naturally occurring NEASs; the higher the certainty rating, the more projective the prejacent. Figure \ref{f-corpus}, which plots the mean certainty ratings of the 59 NEASs by evaluative adjective, shows that the prejacent exhibits projection variability: for some NEASs, the prejacent is highly projective, for some it is weakly projective, and for others it is in-between. Across the 59 NEASs, the prejacent is not highly projective:  the mean certainty rating was only 3.2.\footnote{In contrast to the prejacent of EASs like {\em Kim was smart to watch the movie}, the prejacent of sentences like {\em It was smart of Kim to watch the movie} appears to be more highly projective according to native speaker intuitions, and, therefore, may be best analyzed as a lexically specified presupposition, as in \citealt{oshima09b}.} Details on this corpus-based web study are given in Appendix \ref{a-corpus}.

@@

In the next section, we build on \citepos{best-question} question-based analysis of projective content to develop an analysis that derives the projectivity of the prejacent and the generalization from their status as not-at-issue with respect to the question addressed by the EAS. Section \ref{s3} provides experimental evidence for two predictions of the analysis. After briefly considering the aforementioned interspeaker variation in section \ref{s5}, the paper concludes in section \ref{s6}.

In this section, we develop a question-based projection analysis for EASs. We start in section \ref{s21} by introducing the question-based analysis developed in \citealt{best-question} for the content of the complement of factive predicates. In section \ref{s22}, we extend this analysis to EASs and show how the  interaction between the projectivity of the prejacent and the generalization falls out of the analysis. In section \ref{s23}, we extend the analysis by incorporating a constraint on the questions addressed by utterances of EASs.

@@

The content of the complement of factive predicates like {\em know} or {\em discover} has long been observed to be projective (e.g., \citealt{kiparsky-kiparsky70,karttunen71b}). Thus, the content of the complement of {\em discover} can be understood as a commitment of the speaker/writer, as  in (\ref{discover}a), but it does not have to be, as in (\ref{discover}b). See also the discussion around example (\ref{stop}) with {\em know} in section \ref{s1}.

\begin{exe}
\ex\label{discover}
\begin{xlist}
\ex Caroline stepped forward after a moment. ``There is, of course, also the issue of Alex and the other Enhanced to tackle. If anyone {\bf discovers} that Alex is here, it will be a disaster difficult to avert.'' Alex frowned. \hfill L.C.\ Mawson, {\em Pandora: Freya Snow}, \#13

\ex {[}Mattress springs] also work well to deter rabbits \& foxes from digging into the chook-pen (Hen-run). Dig a shallow trench the width of a single mattress, then place the springs flat into the trench. Drive your fence posts in the mid-line, so half the spring is outside \& half inside the pen. I haven't tried this with wombats, though \& if anyone {\bf discovers} that the method is also wombat-proof, I'd really like to know. \hfill (\citealt[79]{beaver-belly})

Traditionally, projection of the content of the complement has been derived by specifying the content as presupposed in the lexical entry of the factive predicate (e.g., \citealt{heim83,vds92}); as discussed in section \ref{s1}, non-projection under such analyses is attributed to local accommodation. In view of the observation that the projectivity of the content of the complement is weaker than that of other projective content, some analyses have abandoned the assumption that the content projects because it is lexically specified as presupposed: on some analyses, projection is derived from a lexically specified set of alternatives to the factive predicate in combination with pragmatic principles (e.g., \citealt{abusch02,abusch10,romoli2015}) and on others it is derived from the discourse status of the content of the complement (e.g., \citealt{abrusan2011,abrusan2016,brst-salt10,best-question}). Analyses that derive the projection of the content of the complement from its discourse status are empirically motivated by the observation that projection is sensitive to information structure. To illustrate, consider the examples in (\ref{beaver}): \citet{beaver-belly} hypothesized that the content of the complement of {\em discover} is less projective in (\ref{beaver}a), which is realized with focus on {\em plagiarized}, than in (\ref{beaver}b), which is realized with focus on {\em discover}. For experimental evidence for this hypothesis see \citealt{cummins-rohde2015,tonhauser-salt26,djaerv-bacovcin-salt27} and \citealt{mahler-nels}.

\begin{exe}
\ex\label{beaver}
\begin{xlist}
\ex If the T.A.\ discovers that your work is [plagiarized]$_{\mbox{F}}$, I will be forced to notify the Dean.

\ex If the T.A.\  [discover]$_{\mbox{F}}$ that your work is plagiarized, I will be forced to notify the Dean. \\ \hspace*{.2cm} \hfill (adapted from \citealt[93]{beaver-belly})

To capture the information-structure sensitivity of the projection of the content of the complement,  \citealt{brst-salt10,best-question} assumed that the content of the complement is a lexical entailment of sentences with factive predicates that projects if it is not at-issue with respect to the Discourse Question addressed by the utterance of the sentence with the factive predicate.\footnote{Following \citealt{abrusan2011}, lexical entailments are entailments of sentences, not predicates. They differ from logical entailments, which can be derived independently from the meaning of the sentence. For instance, the disjunction of the content of the complement and the proposition that it is raining is a logical entailment of a sentence with a factive predicate, but not a lexical one. For the assumption that at least some projective contents, including presuppositions, are entailments that may project see, e.g., \citealt{gazdar79b,barker02,schlenker10,abrusan2011,abrusan2016} and \citealt{anand-hacquard2014}.} The Discourse Question ``provides the topic of a segment of discourse and imposes relevance constraints on conversational contributions'' (\citealt[192]{best-question}). Utterance content is at-issue with respect to the Discourse Question of the utterance if the content addresses the Discourse Question, i.e., entails at least a partial answer to the question (\citealt{roberts12}/1996). To illustrate, consider the content of the complement of {\em discover} in B's utterances in (\ref{discover2}a) and (\ref{discover2}b), that Harriet was at Princeton for a job interview. In the these examples, the Discourse Questions that B's utterances address are made explicit by A's interrogative utterances. In (\ref{discover2}a), the content of the  complement of {\em discover} is at-issue because it addresses the Discourse Question: that Harriet was at Princeton for a job interview is an answer A's interrogative utterance of where Harriet was yesterday. In (\ref{discover2}b), on the other hand, the content of the complement is not at-issue because it is not an answer to A's interrogative utterance; rather, here the Discourse Question is addressed by the main clause content of B's utterance.

@@

Now consider the example in (\ref{discover22}), in which  {\em discover} is embedded in a polar question: consequently, the speaker is not taken to be committed to the content of the main clause; rather, the speaker is asking whether Henry discovered the content of the complement.

\begin{exe}
\ex\label{discover22} Did Henry discover that Harriet was at Princeton for a job interview?
\end{exe}
In contrast to the main clause content, the content of the complement of {\em discover} may project out of the polar question, as noted above. According to \citealt{best-question}, one of the conditions under which the content of the complement projects is when it is not at-issue with respect to the Discourse Question addressed by the utterance. Consider the examples in (\ref{discover3}), where A's interrogative utterances again make explicit the Discourse Questions addressed by B's utterances. In (\ref{discover3}a), the content of the complement of B's utterance does not address the Discourse Question, i.e., is not at-issue. Here, B is taken to be committed to the content of the complement: to make sense of B's utterance as an answer to A's question, there must be a connection between the possibility of Henry discovering something about Harriet and Henry's mood; an obvious connection is that B takes the content of the complement to be true, i.e., that it is possible that Henry discovered that Harriet was at Princeton for a job interview. Thus, the content of the complement projects. In (\ref{discover3}b), by contrast, the content of the complement addresses the Discourse Question, i.e., is at-issue. Here, B need not be taken to be committed to the content of the complement to make sense of how B's utterance addresses A's question and the content of the complement does not project.

@@

\end{exe}}
